---
title: 'STAT 140: Homework 5'
author: "YOUR NAME HERE"
date: "10/23/2019"
output: html_document
---

## BOOK EXERCISES (OpenIntro Statistics, Fourth Edition)

4.1) **Area under the curve, Part I.** What percent of a standard normal distribution $N(\mu=0,\sigma=1)$ is found in each region? For each part, draw a picture, use appropriate probability notation, and show all your work.

a) Z < -1.35

**Draw a picture.**

**Answer.**

b) Z > 1.48

**Draw a picture.**

**Answer.**

c) -0.4 < Z < 1.5

**Draw a picture.**

**Answer.**

d) |Z| > 2 (|Z| means the absolute value of Z. We can rewrite this question as Z > 2 or Z < -2.)

**Draw a picture.**

**Answer.**

4.2) **Area under the curve, Part II.** What percent of a standard normal distribution is found in each region? For each part, draw a picture, use appropriate probability notation, and show all your work.

a) Z < -1.13

**Draw a picture.**

**Answer.**

b) Z < 0.18

**Draw a picture.**

**Answer.**

c) Z > 8

**Draw a picture.**

**Answer.**

d) |Z| < 0.5

**Draw a picture.**

**Answer.**

4.7) **LA weather, Part I.** The average daily high temperature in June in LA is 77 degrees Fahrenheit with a standard deviation of 5 degrees. Suppose that the temperatures in June closely follow a normal distribution.

a) What is the probability of observing an 83 degree Fahrenheit temperature or higher in LA during a randomly chosen day in June? Draw a picture, use appropriate probability notation, and show all your work.

**Draw a picture:**

**Answer:**

b) How cool are the coldest 10% of the days (days with the lowest average high temperature) during June in LA? Draw a picture, use appropriate probability notation, and show all your work.

**Draw a picture:**

**Answer:**

4.10) **Find the SD.** Cholesterol levels for women aged 20 to 34 follow an approximately normal distribution with mean 185 milligrams per decileter (mg/dl). Women with cholesterol levels above 220 mg/dl are considered to have high cholesterol and about 18.5% of women fall into this category. What is the standard deviation of the distribution of cholesterol levels for women aged 20 to 34? Draw a picture to help summarize what you know, use appropriate notation, and show all your work.

**Draw a picture:**

**Answer:**


4.15) **Bernoulli, the mean.** Use the probability rules from Section 3.4 to derive the mean of a Bernoulli random variable, i.e. a random variable X that takes value 1 with probability p and value 0 with probability 1-p. That is, compute the expected value of a generic Bernoulli random variable.

**Answer:**

4.16) **Bernoulli, the standard deviation.** Use the probability rules from Section 3.4 to derive the standard deviation of a Bernoulli random variable, i.e. a random variable X that takes value 1 with probability p and value 0 with probability 1-p. That is, compute the square root of the variance of a generic Bernoulli random variable.

**Answer:**

4.17) **Underage drinking, Part I.** Data collected by the Substance Abuse and Mental Health Services Administration (SAMSHA) suggests that 69.7% of 18-20 years olds consumed alcoholic beverages in any given year. For the following, when necessary, use appropriate probability notation and show all your work. **Please do these calculations using the binomial formula.** If you wish, you can check your answers using R.

a) Suppose a random sample of ten 18-20 year olds is taken. Is the use of the binomial distribution appropriate for calculating the probability that exactly six consumed alcoholic beverages? Explain. (*Hint: check ALL the Binomial assumptions!*)

**Answer:**

b) Calculate the probability that exactly 6 out of 10 randomly sampled 18-20 year olds consumed an alcoholic drink.

**Answer:**

c) What is the probability that exactly four out of ten 18-20 year olds have *not* consumed an alcoholic beverage?

**Answer:**

d) What is the probability that at most 2 out of 5 randomly sampled 18-20 year olds have consumed alcoholic beverages?

**Answer:**

e) What is the probability that at least 1 out of 5 randomly sampled 18-20 year olds have consumed alcoholic beverages?

**Answer:**

4.35) **Roulette winnings.** In the game of roulette, a wheel is spun and you place bets on where it will stop. One popular bet is that it will stop on a red slot; such a bet has an 18/38 chance of winning. If it stops on red, you double the money you bet. If not, you lose the money you bet. Suppose you play 3 times, each time with a $1 bet. 

a) Let the random variable X represent the number of wins. Does X follow a binomial distribution (check conditions)? What are the parameters?

**Answer:**

b) There are four possible outcomes we can observe. Denote each of the four outcomes using probability notation and find the probability of each.

**Answer:**

**Answer:**

**Answer:**

**Answer:**

c) Let the random variable Y represent the total amount won or lost. Write a probability model for Y (use a probability distribution table like in Chapter 3).

**Answer:**

## R EXERCISES

**Exercise 1.**

Pizza delivery times are approximately normally distributed, with mean 30 minutes and standard deviation 5 minutes. In the R chunk below, 500 pizza delivery times are simulated from a normal distribution with that mean and variance. You will use this simulated data set, pizza_wait_sim, throughout this exercise. This exercise involves as review of the pnorm() and qnorm() functions we saw in lecture.

```{r pizza delivery waiting times}
## Simulated waiting times for pizza delivery
pizza_wait_sim <- rnorm(n=500, mean=30, sd=5)
```

What is the probability that delivery takes less than 20 minutes?

**Answer:**

What is the probability that delivery takes more than 40 minutes? 

**Answer:**

What do you notice about the previous two answers? Why is this happening? (No R required.)

**Answer:**

The fastest 15% of deliveries occur in under how many minutes?

**Answer:**

Half of the deliveries take more than how many minutes?

**Answer:**

Did you need R to answer the previous question? Why or why not?

**Exercise 2.** 

Suppose you are really hungry and only keep track of whether pizza is delivered in less than 20 minutes (or not). We can take the pizza_wait_sim data from above and turn it into a vector that only contains this information and store is as pizza_wait_sim_bin. This exercise involves a review of the pbinom(), dbinom(), and rbinom() functions we saw in class.

```{r pizza delivery waiting times binary}
pizza_wait_sim_bin <- pizza_wait_sim < 20
```

What is the form of this vector (i.e. what is in it)?

**Answer:**

What happens if you apply the sum function to pizza_wait_sim_bin? What does this return and why?

**Answer:**

The probability of a pizza delivery in less than 20 minutes is 0.0228. 

Suppose we randomly sample 10 deliveries. What is the probability of observing no pizza delivery times of less than 20 minutes?

**Answer:**

Suppose we randomly sample 100 deliveries. What is the probabiliy of observing more than 5 pizza delivery times of less than 20 minutes?

**Answer:**

Suppose we randomly sample 100 deliveries. What is the probabiliy of observing exactly 5 pizza delivery times of less than 20 minutes?

**Answer:**

Using the rbinom() and hist() functions, generate and make histograms of 1000 realizations of (i) 10 deliveries, (ii) 100 deliveries, and (iii) 500 deliveries. 

**Answer:**

For (i)-(iii), do any of your histograms look normal? The *success failure condition* states that generally when np > 10 and np(1-p) > 10, it is okay to use the normal approximation to the binomial. Are these conditions met for (i)-(iii)? If so, are they met when your histograms look normal?

**Answer:**

 