---
title: 'Class Activity: Introduction to Missing Data'
subtitle: 'STAT 340: Applied Regression'
output: pdf_document
---

For this activity, consider this data set on chemotherapy for Stage B/C colon cancer, available as `colon_s` from the `finalfit` package in R. The help page for this data set gives the following description: 

"These are data from one of the first successful trials of adjuvant chemotherapy for colon cancer. Levamisole is a low-toxicity compound previously used to treat worm infestations in animals; 5-FU is a moderately toxic (as these things go) chemotherapy agent. There are two records per person, one for recurrence and one for death."



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(finalfit)
```

## Scenario 1

Consider the following pairs plots. Decide whether the data appear to be MCAR, MAR, or MNAR. 

```{r fig.align='center', echo=FALSE, warning=FALSE}
# Create some extra missing data
## Smoking missing completely at random
set.seed(1)
colon_s$smoking = 
  sample(c("Smoker", "Non-smoker", NA), 
    dim(colon_s)[1], replace=TRUE, 
    prob = c(0.2, 0.7, 0.1)) %>% 
  factor() %>% 
    ff_label("Smoking")

explanatory = c("age", "sex.factor", 
  "nodes", "obstruct.factor",  
  "smoking")
dependent = "mort_5yr"
colon_s %>% 
  missing_pairs(dependent, explanatory)
```


## Scenario 2

```{r fig.align='center', echo=FALSE, warning=FALSE}
## Smoking missing conditional on patient sex
colon_s$smoking[colon_s$sex.factor == "Female"] = 
  sample(c("Smoker", "Non-smoker", NA), 
    sum(colon_s$sex.factor == "Female"), 
    replace = TRUE,
    prob = c(0.1, 0.5, 0.4))

colon_s$smoking[colon_s$sex.factor == "Male"] = 
  sample(c("Smoker", "Non-smoker", NA), 
    sum(colon_s$sex.factor == "Male"), 
    replace=TRUE, prob = c(0.15, 0.75, 0.1))
colon_s$smoking = factor(colon_s$smoking) %>% 
    ff_label("Smoking")

explanatory = c("age", "sex.factor", 
  "nodes", "obstruct.factor",  
  "smoking")
dependent = "mort_5yr"
colon_s %>% 
  missing_pairs(dependent, explanatory)
```

Consider the following pairs plots. Decide whether the data appear to be MCAR, MAR, or MNAR. Support your conclusions with observations from the pairs plot.