---
title: "R Matrix Commands"
output: pdf_document
---

## R commands

- c(): create a vector
- cbind(): bind vectors together as columns of a matrix
- t(): transpose of a matrix
- %*%: matrix multiplication
- solve(): find the inverse of a matrix
- matrix(): directly create a matrix from specified values; by default this will make a matrix with 1 column

## Examples

### Create model matrix: $\mathbf{X}$

```{r}
X <- cbind(
  c(1,1,1),
  c(0,1,1)
)
X
```

### Matrix multiplication: $\mathbf{X'X}$

```{r}
t(X)%*%X
```

### Matrix inversion: $(\mathbf{X'X})^{-1}$

```{r}
solve(t(X)%*%X)
```

### Create $\mathbf{y}$

```{r}
# suppose y_1=2, y_2=4, y_3=6
y <- matrix(c(1,2,3))
```

### Helpful side note (may be useful in the future)

```{r}
example <- matrix(1:6, nrow=2, ncol=3)
example
```

### $\hat{\boldsymbol{\beta}}$ computed using matrices

```{r}
beta_hat <- solve(t(X)%*%X) %*% t(X) %*% y
```

### Compute hat matrix (also called the projection matrix)

```{r}
H <- X %*% solve(t(X)%*%X) %*% t(X)
H
```

### Compute $\hat{\mathbf{y}}$

$\hat{\mathbf{y}}=\mathbf{X}\hat{\boldsymbol{\beta}}$

```{r}
y_hat1 <- X %*% beta_hat
y_hat1
```

$\hat{\mathbf{y}}=\mathbf{H}\mathbf{y}$

```{r}
y_hat2 <- H %*% y
y_hat2
```

### Compute residuals

$\hat{\boldsymbol{\epsilon}}=\mathbf{y}-\hat{\mathbf{y}}$

```{r}
resids <- y-y_hat1
resids
```

### Recall, all of this came from minimizing the residual sum of squares (RSS)!
```{r}
rss <- sum(resids^2)
rss
```


